---
title: "StatsLab Code"
author:
- Justin Glommen (Scenarios 1,4)
- Victor Angulo (Scenarios 3,6)
- Atharva Fulay (Scenario 2)
- Peter Yao (Scenario 5)
date: "2/5/2017"
output:
  pdf_document: default
  word_document: default
---

# Data Management
Loading data from current directory
```{r}
data <- read.table("videodata.txt", header=TRUE)
data.population <- 314     # True population
data.samples <- 91         # Number of samples
head(data)
summary(data)
```


## Cleaning Data
Replacing 99 values (the unanswered/improper results) with NAs
```{r}
data[data == 99] <- NA
numSamples <- NROW(data)
head(data)
summary(data)
```



# Scenario 1

## Sample Proportion of Students Who Played a Video Game in the Last Week
The individual variables measured here are Bernoulli since time is being converted to a binary 'did' or 'did not' play.
```{r}
# Create 'numPlayers' variable to count number of players in the last week.
# This is done by counting the number of people with time spent over 0, which represents the
# people who played something in the last week since they spent time on it. 0 indicates no time 
# spent.
numPlayers <- NROW(which(data$time > 0))
paste("Number of players:", numPlayers, sep=" ")

# Sample proportion is the ratio of numPlayers to total students (rows in data)
data.playersSampleProportion <- (numPlayers/numSamples)
paste("Sample proportion:", data.playersSampleProportion, sep=" ")

```


## Players Sample Proportion Confidence Interval
Since the sample Bernoulli variables are NOT identically independentally distributed, the confidence interval itself will be computed utilizing the finite population correction factor.
```{r}
# Sample proportion is nearly Binomial, except not iid.
playersCorrectionFactor <- sqrt((data.population - numSamples)/data.population)
# Binomial standard error formula without correction
playersIndepStandardError <-  (sqrt(data.playersSampleProportion*(1-data.playersSampleProportion))/sqrt(numSamples - 1))
# Standard error with finite population correction
data.playersStandardErrorEstimate <- playersIndepStandardError*playersCorrectionFactor
paste("Corrected Standard Error:", data.playersStandardErrorEstimate, sep=" ")

# Since the sample proportion follows a normal distribution by the Central Limit Theorem,
# we need to multiply the corrected standard error by 1.96 to generate the interval.
data.playersMarginOfError <- 1.96*data.playersStandardErrorEstimate
paste("Margin of Error: ", data.playersMarginOfError, sep="")

# Therefore, the confidence interval:
playersLowerBound <- data.playersSampleProportion - data.playersMarginOfError
playersUpperBound <- data.playersSampleProportion + data.playersMarginOfError
data.playersSampleProportionConf95 <- c(playersLowerBound, playersUpperBound)
paste("Player Proportion 95% CI: ", "(",playersLowerBound, ", ", playersUpperBound,")", sep="")
```


# Scenario 2
```{r}
smalltime.ind <- which(data$time < 6)
data.smalltime <- data[smalltime.ind,]

zerohours.ind <- which(data.smalltime$time ==0)
data.zerohours <- data[zerohours.ind, ]
mean(data.zerohours$freq, na.rm=TRUE)

fewhours.ind <- which(data.smalltime$time > 0 & data.smalltime$time <=5 )
data.fewhours <- data[fewhours.ind, ]
mean(data.fewhours$freq, na.rm=TRUE)

manyhours.ind <- which(data$time > 6)
data.manyhours <- data[manyhours.ind, ]
summary(data.manyhours$freq, na.rm=TRUE)

daily.ind <- which(data$freq == 1)
weekly.ind <- which(data$freq == 2)
monthly.ind <- which(data$freq == 3)
semester.ind <- which(data$freq == 4)

data.daily <- data[daily.ind, ]
data.weekly <- data[weekly.ind, ]
data.monthly <- data[monthly.ind, ]
data.semester <- data[semester.ind, ]

mean(data.daily$time)
mean(data.weekly$time)
mean(data.monthly$time)
mean(data.semester$time)

busy.ind <- which(data$busy == 1)
data.busy <- data[busy.ind, ]

notbusy.ind <- which(data$busy == 0)
data.notbusy <- data[notbusy.ind, ]

mean(data.busy$time)
mean(data.notbusy$time)
```



# Scenario 3
```{r}
#First we calculate the estimate for the # of students that played a video game:
nogame.ind <- which(data['time'] == 0.0) #Identify those who did not play video games the week prior
data.nogame <- data[nogame.ind,] #Create a data frame with no gamers
n1 <- length(data.nogame$time) #Calculates the # of students that played video games
prop.nogame <- (n1)/91 #Calculates the proportion (# that don't play/sample size)
sd.prop.nogame <- sqrt( (.6263736)*(1-.6263736)/90 )*sqrt((314-91)/314 ) #Calculates the sd of those that don't game
prop.nogame.ci <- prop.nogame + c(-1, 1)*2*sd.prop.nogame #Creates the CI

#Histogram of sample time spent playing
hist(data$time, main = "Histogram of Time Spent Playing Videogames", xlab = "Amount of Hours", ylab = "Amount of Students", 
     col = 4, density = 15, breaks = 15)

#Here we do Bootstrap
boot.population <- rep(data$time, length.out = 314) #Creates the population
sample1 <- sample(boot.population, size = 91, replace = FALSE) #creates the sample populations
B = 500 # the number of bootstrap samples we want
boot.sample <- array(dim = c(B, 91))
for (i in 1:B)
  {
  boot.sample[i, ] <- sample(boot.population, size = 91, replace = FALSE)
}
boot.mean <- apply(X = boot.sample, MARGIN = 1, FUN = mean) #Here we take the sample mean of each sample
ci.boot <- c(quantile(boot.mean,0.025), quantile(boot.mean, 0.975))

```


# Scenario 4
Getting proportion who likes games.
```{r}
# Initializing variables corresponding to responses from students on the survey
likeVeryMuch <- 2
likeSomewhat <- 3
# Fetching all students who responded with positive game likeness
data.likeColumns <- which(data$like == likeVeryMuch) 
data.likeColumns <- c(data.likeColumns, which(data$like == likeSomewhat))
# Calculating percentage
numOfLikes <- NROW(data.likeColumns)
proportionLike <- numOfLikes/data.samples
paste("Proportion of Like: ", proportionLike, sep="")
```

# Scenario 5
```{r}
#didn't use graphing or cross tabulation
#18+30=48 people who own a computer like playing games
play2_comp <- NROW(which(data$like==2 & data$own == 1))
play3_comp <- NROW(which(data$like==3 & data$own == 1))

#1+12+5=18 people who own a computer don't like playing games
noplay1_comp <- NROW(which(data$like==1 & data$own==1))
noplay4_comp <- NROW(which(data$like==4 & data$own==1))
noplay5_comp <- NROW(which(data$like==5 & data$own==1))

#5+16=21 people who don't own a computer like playing games
play2_nocomp <- NROW(which(data$like==2 & data$own == 0))
play3_nocomp <- NROW(which(data$like==3 & data$own == 0))

#0+1+2=3 people who don't own a computer don't like playing games
noplay1_nocomp <- NROW(which(data$like==1 & data$own==0))
noplay4_nocomp <- NROW(which(data$like==4 & data$own==0))
noplay5_nocomp <- NROW(which(data$like==5 & data$own==0))

#14+25=39 people who like games worked
play2_work <- NROW(which(data$like==2 & data$work>0))
play3_work <- NROW(which(data$like==3 & data$work>0))

#1+3+3=7 people who don't like games worked
play1_work <- NROW(which(data$like==1 & data$work>0))
play4_work <- NROW(which(data$like==4 & data$work>0))
play5_work <- NROW(which(data$like==5 & data$work>0))

#9+21=30 people who like games don't work
play2_nowork <- NROW(which(data$like==2 & data$work==0))
play3_nowork <- NROW(which(data$like==3 & data$work==0))

#0+10+4=14 people who don't like games don't worked
noplay1_nowork <- NROW(which(data$like==1 & data$work==0))
noplay4_nowork <- NROW(which(data$like==4 & data$work==0))
noplay5_nowork <- NROW(which(data$like==5 & data$work==0))

#18+25=43 males who like games
play2_male <- NROW(which(data$like==2 & data$sex==1))
play3_male <- NROW(which(data$like==3 & data$sex==1))

#1+5+3=9 males who don't like games
play1_male <- NROW(which(data$like==1 & data$sex==1))
play4_male <- NROW(which(data$like==4 & data$sex==1))
play5_male <- NROW(which(data$like==5 & data$sex==1))

#5+21=26 females who like games
play2_female <- NROW(which(data$like==2 & data$sex==0))
play3_female <- NROW(which(data$like==3 & data$sex==0))

#0+8+4=12 females who don't like games
play1_female <- NROW(which(data$like==1 & data$sex==0))
play4_female <- NROW(which(data$like==4 & data$sex==0))
play5_female <- NROW(which(data$like==5 & data$sex==0))
```

# Scenario 6
```{r}
#Chi-square test
observed <- c(31, 52, 8, 0)
expected <- c(.2, .33, .4, .1)
chisq.test(observed, p = expected, rescale.p = TRUE)
```
